---
title: "Chapter 03 Sampling the imaginary"
author: "Adam Gruer"
date: "22/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What I remember




## Notes


## Questions

```{r}

p_grid <- seq( from=0 , to=1 , length.out=1000 )
prob_p <- rep( 1 , 1000 )
prob_data <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)

samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
plot(samples)

library(rethinking)
dens( samples )

# posterior probability prop water below 0.5
sum(posterior[p_grid < 0.5])

sum(samples < 0.5) / 1e4
sum(samples > 0.5 & samples < 0.75) / 1e4

# compatibility interval
quantile(samples, 0.8)

quantile(samples, c(0.1, 0.9))

# observed w, w, w
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep(1,1000)
likelihood <- dbinom( 3 , size=3 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
samples <- sample( p_grid , size=1e4 , replace=TRUE , prob=posterior )
dens(samples)

PI(samples,prob =  0.5)
HPDI(samples, prob = 0.5)
p_grid[which.max(posterior)]

chainmode(samples, adj = 0.01)

mean(samples)
median(samples)
```
## Simulaiton

```{r}
dbinom(0:2, size = 2, prob = 0.7)

rbinom(1, size = 2, prob = 0.7)

rbinom(10, size = 2, prob = 0.7)

dummy_w <- rbinom(1e5, size = 2, prob = 0.7)
table(dummy_w) /1e5

dummy_w <- rbinom( 1e5 , size=9 , prob=0.7 )
table(dummy_w) /1e5
simplehist( dummy_w , xlab="dummy water count" )
```

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prob_p <- rep( 1 , 1000 )
prob_data <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)

samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )

w <- rbinom(1e4, size = 9, prob = 0.6)
simplehist(w)
w <- rbinom(1e4, size = 9, prob = samples)
simplehist(w)
```

## practice

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
dens(samples)
```
## 3E1

```{r}
sum(samples < 0.2) / 1e4
```
## 3E2

```{r}
sum(samples > 0.8) / 1e4
```
## 3E3

```{r}
sum(samples > 0.2 & samples < 0.8) / 1e4
```
## 3E4

```{r}
quantile(samples , 0.2)
```

## 3E5

```{r}
quantile(samples, 0.8)
```

## 3E6

```{r}
rethinking::HPDI(samples, prob = 0.66)

```

## 3E7

```{r}
PI(samples, prob = 0.66)
quantile(samples, c(0.17, 0.83))
```

## 3M1

Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as before.

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

```

## 3M2

Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HPDI for p

```{r}
set.seed(100)
samples <- sample(p_grid, 1e4, replace = TRUE, prob = posterior)
HPDI(samples, 0.9)
```

## 3M3

Construct a posterior predictive check for this model and data. This means simulate the distribution of samples, averaging over the posterior uncertainty in p. What is the probability of observing
8 water in 15 tosses?

```{r}
set.seed(100)
w <- rbinom(1e4, size = 15, prob = samples)
mean(w == 8)
simplehist(w)
```

## 3M4

Using the posterior distribution constructed from the new (8/15) data, now calculate the probability of observing 6 water in 9 tosses.

```{r}
set.seed(100)
w <- rbinom(1e4, size = 9, prob = samples)
simplehist(w)
mean(w == 6)
```

## 3M5.

Start over at 3M1, but now use a prior that is zero below p = 0.5 and a constant above p = 0.5.
This corresponds to prior information that a majority of the Earth’s surface is water. Repeat each
problem above and compare the inferences. What difference does the better prior make? If it helps,
compare inferences (using both priors) to the true value p = 0.7.
```{r}

prior2 <- ifelse(p_grid < 0.5, 0, 1)
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior2 <- likelihood * prior2
posterior2 <- posterior2 / sum(posterior2)

set.seed(100)
samples2 <- sample(p_grid, 1e4,  replace = TRUE, prob = posterior2)
dens(samples2)

HPDI(samples2, 0.9)

set.seed(100)
w2 <- rbinom(1e4, size = 15, prob = samples2)
mean(w2 == 8)
simplehist(w2)

set.seed(100)
w2 <- rbinom(1e4, size = 9, prob = samples2)
simplehist(w2)
mean(w2 == 6)

```

Introduction. The practice problems here all use the data below. These data indicate the gender
(male=1, female=0) of officially reported first and second born children in 100 two-child families.

```{r}
library(rethinking)
data(homeworkch3)
```

Use these vectors as data. So for example to compute the total number of boys born across all of these
births, you could use:

```{r}
sum(birth1 + birth2)
```

## 3H1. 

Using grid approximation, compute the posterior distribution for the probability of a birth
being a boy. Assume a uniform prior probability. Which parameter value maximizes the posterior
probability?

```{r}
p_grid = seq(from = 0, to = 1, length.out = 1000)
prior <- rep(0.5, 1000)
set.seed(100)
likelihood <- dbinom(111, size = 200, prob = p_grid )

posterior <- prior * likelihood 
posterior <- posterior / sum(posterior)

plot(p_grid, posterior)
p_grid[which.max(posterior)]
```

## 3H2

Using the sample function, draw 10,000 random parameter values from the posterior distribution you calculated above. Use these samples to estimate the 50%, 89%, and 97% highest posterior
density intervals.

```{r}
set.seed(100)
samples <- sample(p_grid, size = 1e4, replace = TRUE,  prob = posterior )

HPDI(samples, prob = 0.5)
HPDI(samples, prob = 0.89)
HPDI(samples, prob = .97)
```

## 3H3.

Use rbinom to simulate 10,000 replicates of 200 births. You should end up with 10,000 numbers, each one a count of boys out of 200 births. Compare the distribution of predicted numbers
of boys to the actual count in the data (111 boys out of 200 births). There are many good ways to
visualize the simulations, but the dens command (part of the rethinking package) is probably the
easiest way in this case. Does it look like the model fits the data well? That is, does the distribution
of predictions include the actual observation as a central, likely outcome?

Yes

```{r}
set.seed(100)
b <- rbinom(1e4, size = 200, prob = samples) 

dens(b)
```

##3H4.

Now compare 10,000 counts of boys from 100 simulated first borns only to the number of boys
in the first births, birth1. How does the model look in this light?


No the model tends to overestimate the number of first born boys. 

```{r}
set.seed(100)
b100 <- rbinom(1e4, size = 100, prob = samples)
dens(b100)
library(ggplot2)
ggplot(data = tibble::tibble(nboys = b100)) +
  geom_density(aes(x = nboys), adjust = 0.5) +
  geom_vline(xintercept  = sum(birth1))
dens(b100)
```

## 3H5. 

The model assumes that sex of first and second births are independent. To check this assumption, focus now on second births that followed female first borns. Compare 10,000 simulated counts
of boys to only those second births that followed girls. To do this correctly, you need to count the
number of first borns who were girls and simulate that many births, 10,000 times. Compare the
counts of boys in your simulations to the actual observed count of boys following girls. How does the
model look in this light? Any guesses what is going on in these data?

The model significantly underestimates the number of second children who are boys following a first born girl.  Perhaps indicates that sex of first and second born children are not independant of each other.

```{r}
g1 <- 100 - sum(birth1)
boys_after_girl<- rbinom(1e4, size = g1, prob = samples)

ggplot(data = tibble::tibble(nboys = boys_after_girl)) +
  geom_density(aes(x = nboys), adjust = 0.8) +
  geom_vline(xintercept  = sum(birth2[birth1 == 0]))

```

