---
title: "chapter 4 practice"
author: "Adam Gruer"
date: "02/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
library(dplyr)
library(tidyr)
library(ggplot2)
```


# Easy

## 4E1

y ~ Normal(mu, sigma) is the likelihood. 
The likelihood shows the distribution
of the variable being predicted

## 4E2 

There are two parameters in the posteriod distribution, mu and sigma

## 4E3

mu = a + b * x is the linear model

## 4E4



3 parameters a, b and sigma . mu is no longer a parameter as it is now a fnciton
of the other parameters

# Medium

## 4M1

y ~ Normal(mu, sigma)
mu ~ Normal(0, 10)
sigma ~ Exponential(1)

Simulate observed y from the prior
```{r}
sample_mu <- rnorm(1e4, 0, 10)
sample_sigma <- rexp(1e4, 1)
plot(sample_mu)
dens(sample_mu)
plot(sample_sigma)
dens(sample_sigma)
prior_y <- rnorm(1e4, sample_mu, sample_sigma)
plot(prior_y)
dens(prior_y, adj = 0.9)

```
## 4M2

```{r}
flist <- alist(
    y ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 10),
    sigma ~ exp(1)
  )

```

## 4M3

translate quap model formula to mathematical model definition

```{r}
flist <- alist(
  y ~ dnorm(mu, sigma),
  mu <- a + b*x,
  a ~ dnorm(0, 10),
  b ~ dunif(0, 1),
  sigma ~ dexp(1)
)


```

y_i ~ Normal(mu_i, sigma)
mu_i = a + b * x_i
a ~ Normal(0, 10)
b ~ Uniform(0,1)
sigma ~ Exponential(1)

## 4M4

h_i ~ Normal(mu_i, sigma)
mu_i = a + b * (x - xbar)
a ~ Normal(130, 10)
b ~ LogNormal(0, 1) 
sigma ~ Exponential(1)

Assuming primary students . b 
```{r}
n <- 1000
tibble(group = seq_len(n),
       alpha = rnorm(n, 130, 10),
       beta = rlnorm(n, 0, 1 )) %>%
  expand(nesting(group, alpha, beta), year = c(1, 3)) %>%
  mutate(height = alpha + beta * (year - 2)) %>%
  ggplot(aes(x = year, y = height, group = group)) +
  geom_line(alpha = 1/10) +
  geom_hline(yintercept = c(0, 272), linetype = 2:1, color = "red") +
  annotate(geom = "text", x = 1, y = 0, hjust = 0, vjust = 1,
           label = "Embryo") +
  annotate(geom = "text", x = 1, y = 272, hjust = 0, vjust = 0,
           label = "World's tallest person (272cm)") +
  coord_cartesian(ylim = c(-25, 300)) +
  labs(x = "Year", y = "Height")

```
## 4M5

Each child gets taller each year
I don't understand log scales well enough 
in my ansewer above I use rlnorm(0,1)
in the  online solutions rlnorm(2, 0.5) is used . i dont understand why https://sr2-solutions.wjakethompson.com/linear-models.html#chapter-4

```{r}
samples_1 <- rlnorm(1e4, 0, 1)
PI_1 <- PI(samples_1, prob = 0.89)
PI_1
dens(samples_1)

samples_2 <- rlnorm(1e4, 2,0.5)
PI_2 <- PI(samples_2, prob = 0.89)
PI_2
```
Actually I prefer my choice of prior based on experience with my own children.
Choice definitely depends on starting age of children


## 4M6


# Hard

## 4H1

Make predictions

```{r}
individuals <- c(46.95, 43.72, 64.78, 32.59, 54.63)

data("Howell1")
d <- Howell1
d2 <- d[d$age >= 18, ]


xbar <- mean(d2$weight)

#fit omdel
m4.3 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * (weight - xbar),
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0, 50)
  ),
  data = d2
)

post <- extract.samples(m4.3)
mu <- link(m4.3, data = data.frame(weight = individuals))
```

```{r}
mu.df <- `colnames<-`(mu, individuals) %>% 
  as.data.frame() %>% 
  pivot_longer(cols = everything(),
               names_to = "weight",
               values_to = "height",
               names_transform = list(weight = as.numeric)) 

ggplot(data = mu.df, aes(x = weight, y = height)) +
  geom_point(alpha = 0.2 )
 
```


```{r}
#summarise distribution of mu
mu.df <- mu.df %>% 
  group_by(weight) %>% 
  summarise(expected_height = mean(height),
            int = PI(height, prob = 0.89),
            probs= c("5%", "94%"), 
            .groups = "drop") %>%  
  pivot_wider( id_cols = c(weight, expected_height), 
               names_from = probs, values_from = int)

```

## 4H2

```{r}
d3 <- d[d$age < 18, ]

xbar <- mean(d3$weight)


youth <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * (weight - xbar),
    a ~ dnorm(80, 40),
    b ~ dlnorm(2, 0.5),
    sigma ~ dunif(0, 50)
  ),
  data = d3
)

precis(youth)
```

For every 10 units of increase in weight this model predicts a child gets 27cm taller

```{r}

post <- extract.samples(youth)

a_map <- mean(post$a)
b_map <- mean(post$b)


weight.seq <- seq(from = 4, to = 45, by = 1)
mu <- link(youth, data = data.frame(weight=weight.seq))

mu.PI <- apply(mu, 2, PI, prob = 0.89)
mu.PI.df <- PI.df(mu.PI, weight.seq)

sim.height <- sim(youth, data = list(weight = weight.seq), n = 1e4)

height.PI <- apply(sim.height, 2, PI , prob= 0.89)
height.PI.df <- PI.df(height.PI, weight.seq)


ggplot() +
  
  geom_function(fun = function(x) {a_map + b_map*(x - xbar)}) +
  geom_ribbon(data = mu.PI.df, 
              aes(x = x, ymin = x5_percent, ymax = x94_percent),
              alpha = 0.2, colour = "grey90"  ) +
   geom_ribbon(data = height.PI.df, 
              aes(x = x, ymin = x5_percent, ymax = x94_percent),
              alpha = 0.2, colour = "grey80"  ) +
  geom_point(data = d3, aes(x = weight, y = height)) 
  
```
Model doesnt do a great job of predicting hieght in infants or late teens.
Maybe do a polynomial linear regression intead


## 4H3

```{r}
m4h3 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * log(weight),
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0, 50)
  ),
  data = d
)

precis(m4h3)
```

```{r}

post <- extract.samples(m4h3)
a_map <- mean(post$a)
b_map <- mean(post$b)

weight.seq <- seq(from = 4, to = 63, by = 1)
mu <- link(m4h3, data = data.frame(weight=weight.seq))
mu.PI <- apply(mu, 2, PI, prob = 0.97)
mu.PI.df <- PI.df(mu.PI, weight.seq)

sim.height <- sim(m4h3, data.frame(weight = weight.seq))
height.PI <- apply(sim.height, 2, PI, prob = 0.97)
height.PI.df <- PI.df(height.PI, weight.seq)

ggplot() +
     geom_ribbon(data = height.PI.df, 
              aes(x = x, ymin = x2_percent, ymax = x98_percent),
              alpha = 0.2, colour = "grey80"  ) +
    geom_ribbon(data = mu.PI.df, 
              aes(x = x, ymin = x2_percent, ymax = x98_percent),
              alpha = 0.2, colour = "grey90"  ) +

 geom_point(data = d, aes(x = weight, y = height), alpha = 0.4, shape = 1, col = rangi2) +
  geom_function(fun = function(x) {a_map + b_map * log(x)})
```


## 4H4




